---
title: "Development, Performance and Survival"
output:
  html_document:
    df_print: paged
---


Analyses including only sets with data on all three variables: development, performance and survival.
```{r, include=FALSE}
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
library(effects)
library(ggeffects)
library(Hmisc)
library(phyr)
library(ape)
library(nlme)
library(stringr)
library(phylotools)
library(phytools)
library(rr2)
```


```{r, message = FALSE}
Lowresponses <- read_csv("~/Desktop/Projects/ThermalTManuscriptScripts/LowLimitsDPS.csv")
Highresponses <- read_csv("~/Desktop/Projects/ThermalTManuscriptScripts/HighLimitsDPS.csv")
Breadth <- read_csv("~/Desktop/Projects/ThermalTManuscriptScripts/BreadthDPS.csv")
```

We discarded "eggtoemergence" category in the subsequent analyses because it is a combination of the other three. We also discarded data from sets that were a combination of localities "combination".

```{r}
# convert character to factor
Lowresponses <- Lowresponses %>%
  mutate_if(is.character, factor)
# discard non-pertinent data

LowLimits <- Lowresponses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa", 
         #quality!= "inferred", 
         quality != "combination") %>% 
  mutate(absLat = abs(lat)) 
LowLimits$lifestage <- factor(LowLimits$lifestage)
LowLimits$sp <- factor(LowLimits$sp)


# High limits
Highresponses <- Highresponses %>%
  mutate_if(is.character, factor)
# discard non-pertinent data

HighLimits <- Highresponses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa", 
         #quality!= "inferred", 
         quality != "combination") %>% 
  mutate(absLat = abs(lat)) 
HighLimits$lifestage <- factor(HighLimits$lifestage)
HighLimits$sp <- factor(HighLimits$sp)

# Breadth

Breadthresponses <- Breadth %>%
  mutate_if(is.character, factor)
# discard non-pertinent data

BreadthLimits <- Breadthresponses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa", 
         #quality!= "inferred", 
         quality != "combination") %>% 
  mutate(absLat = abs(lat)) 
BreadthLimits$lifestage <- factor(BreadthLimits$lifestage)
BreadthLimits$sp <- factor(BreadthLimits$sp)
```

Get how many sets were included in these comparative analysis:
59 sets including 40 species
```{r}
lowsets <- unique(Lowresponses$set)
highsets <- unique(Highresponses$set)
breadthsets <- unique(Breadth$set)

totalsets <- tibble(set = c(lowsets, highsets,breadthsets))
length(unique(totalsets$set)) # we had 59 total sets
lowsp <- unique(Lowresponses$sp)
highsp <- unique(Highresponses$sp)
breadthsp <- unique(Breadth$sp)

totalsp <- tibble(sp = c(lowsp, highsp,breadthsp))
length(unique(totalsp$sp)) # 40 species

```

# Variation of low limits across latitude and lifestages
Development
```{r}

Dl_i <- lmer(Dl ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(LowLimits)) #
Dl_a <- lmer(Dl ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(LowLimits)) #

anova(Dl_i, Dl_a, test.statistic = "Chisq") 
```

```{r}
Dl_a_ML <- lmer(Dl ~ -1 + absLat + lifestage +  (1|sp), REML = F, data = as.data.frame(LowLimits)) #
summary(Dl_a_ML)
#plot(Dl_a_ML)

```

```{r}
# Predictions plot
LowLimits$group <- factor(LowLimits$lifestage)
predlow <- ggpredict(Dl_i, terms = c("absLat","lifestage"))
Dlow_plot <- ggplot(predlow, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line(size = 2 )+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1,linetype = 2)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(D[l]))+
  xlab("")+
  ylim(5, 30)+
  annotate("text", x = 0, y = 30, label = "D", size = 5)+
  #geom_point(data = LowLimits, mapping = aes(x = absLat, y = Dl, col = group), alpha = 0.6, shape = 16)+
  #geom_point(data = LowLimits, mapping = aes(x = absLat, y = Dmin, col = group), alpha = 1, shape = 22)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none", plot.title = element_text(hjust = 0.5))


Dlow_plot
```


Performance

```{r}

Pl_i <- lmer(Pl ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(LowLimits)) #
Pl_a <- lmer(Pl ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(LowLimits)) #

anova(Pl_i, Pl_a, test.statistic = "Chisq") 
```

```{r}
Pl_a_ML <- lmer(Pl ~ -1 + absLat + lifestage +  (1|sp), REML = F, data = as.data.frame(LowLimits)) #
summary(Pl_a_ML)
#plot(Pl_a_ML)
```

```{r}
# Predictions plot
LowLimits$group <- factor(LowLimits$lifestage)
Ppredlow <- ggpredict(Pl_i, terms = c("absLat","lifestage"))
Plow_plot <- ggplot(Ppredlow, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line(size = 2)+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1, linetype = 2)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(P[l]))+
  xlab("")+
  ylim(5, 30)+
  annotate("text", x = 0, y = 30, label = "F", size = 5)+
  #geom_point(data = LowLimits, mapping = aes(x = absLat, y = Pl, col = group), alpha = 0.6, shape = 16)+
  #geom_point(data = LowLimits, mapping = aes(x = absLat, y = Dmin, col = group), alpha = 1, shape = 22)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none", plot.title = element_text(hjust = 0.5))

# Save model predictions for plots
#write_csv(Ppredlow,"~/Desktop/Projects/ThermalTManuscriptScripts/Pl-DP.csv")


Plow_plot
```


Survival

```{r}

Sl_i <- lmer(Sl ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(LowLimits)) #
Sl_a <- lmer(Sl ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(LowLimits)) #

anova(Sl_i, Sl_a, test.statistic = "Chisq") 
```

```{r}
Sl_a_ML <- lmer(Pl ~ -1 + absLat + lifestage +  (1|sp), REML = F, data = as.data.frame(LowLimits)) #
summary(Sl_a_ML)
# plot(Sl_a_ML)
```

```{r}
# Predictions plot
LowLimits$group <- factor(LowLimits$lifestage)
Spredlow <- ggpredict(Sl_i, terms = c("absLat","lifestage"))
Slow_plot <- ggplot(Spredlow, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line(size = 2)+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1, linetype = 2)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(S[l]))+
  xlab("")+
  ylim(5, 30)+
  annotate("text", x = 0, y = 30, label = "E", size = 5)+
  #geom_point(data = LowLimits, mapping = aes(x = absLat, y = Pl, col = group), alpha = 0.6, shape = 16)+
  #geom_point(data = LowLimits, mapping = aes(x = absLat, y = Dmin, col = group), alpha = 1, shape = 22)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")

# Save model predictions for plots
#write_csv(Ppredlow,"~/Desktop/Projects/ThermalTManuscriptScripts/Pl-DP.csv")


Slow_plot
```


lowplots

```{r}
#plot_grid(Dlow_plot, Slow_plot, Plow_plot, ncol = 3)
```



# Variation of High limits accros latitudes and lifestages

```{r}

Dh_i <- lmer(Dh ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(HighLimits)) #
Dh_a <- lmer(Dh ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(HighLimits)) #

anova(Dh_i, Dh_a, test.statistic = "Chisq") 
```

```{r}
Dh_i_ML <- lmer(Dh ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(HighLimits)) #
summary(Dh_i_ML)
#plot(Dh_i_ML)
```

```{r}
# Predictions plot
HighLimits$group <- factor(HighLimits$lifestage)
Dpredhigh <- ggpredict(Dh_i_ML, terms = c("absLat","lifestage"))
Dhigh_plot <- ggplot(Dpredhigh, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line(size = 2)+
 geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1, linetype = 2)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(D[h]))+
  xlab("")+
  ylim(25, 40)+
  annotate("text", x = 0, y = 40, label = "A", size = 5)+
  ggtitle(label = "Development")+
 # geom_point(data = HighLimits, mapping = aes(x = absLat, y = Dh, col = group), alpha = 0.6, shape = 16)+
  #geom_point(data = HighLimits, mapping = aes(x = absLat, y = Dmin, col = group), alpha = 1, shape = 22)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none", plot.title = element_text(hjust = 0.5))



Dhigh_plot
```


Performance

```{r}

Ph_i <- lmer(Ph ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(HighLimits)) #
Ph_a <- lmer(Ph ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(HighLimits)) #

anova(Ph_i, Ph_a, test.statistic = "Chisq") 
```

```{r}
Ph_i_ML <- lmer(Ph ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), REML = F, data = as.data.frame(HighLimits)) #
summary(Ph_i_ML)
#plot(Ph_i_ML)
```

```{r}
# Predictions plot
HighLimits$group <- factor(HighLimits$lifestage)
Ppredhigh <- ggpredict(Ph_i, terms = c("absLat","lifestage"))
Phigh_plot <- ggplot(Ppredhigh, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line(size = 2)+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1, linetype = 2)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(P[h]))+
  xlab("")+
  ylim(25, 40)+
  annotate("text", x = 0, y = 40, label = "C", size = 5)+
  ggtitle(label = "Performance")+
  #geom_point(data = HighLimits, mapping = aes(x = absLat, y = Ph, col = group), alpha = 0.6, shape = 16)+
  #geom_point(data = LowLimits, mapping = aes(x = absLat, y = Dmin, col = group), alpha = 1, shape = 22)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none", plot.title = element_text(hjust = 0.5))

# Save model predictions for plots
#write_csv(Ppredhigh,"~/Desktop/Projects/ThermalTManuscriptScripts/Ph-DP.csv")


Phigh_plot
```


Survival

```{r}

Sh_i <- lmer(Sh ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(HighLimits)) #
Sh_a <- lmer(Sh ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(HighLimits)) #

anova(Sh_i, Sh_a, test.statistic = "Chisq") 
```

```{r}
Sh_i_ML <- lmer(Sh ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), REML = F, data = as.data.frame(HighLimits)) #
summary(Sh_i_ML)
#plot(Sh_i_ML)
```

```{r}
# Predictions plot
HighLimits$group <- factor(HighLimits$lifestage)
Spredhigh <- ggpredict(Sh_i, terms = c("absLat","lifestage"))
Shigh_plot <- ggplot(Spredhigh, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line(size = 2)+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1, linetype = 2)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(S[h]))+
  xlab("")+
  ylim(25, 40)+
  ggtitle(label = "Survival")+
  annotate("text", x = 0, y = 40, label = "B", size = 5)+
  #geom_point(data = HighLimits, mapping = aes(x = absLat, y = Ph, col = group), alpha = 0.6, shape = 16)+
  #geom_point(data = LowLimits, mapping = aes(x = absLat, y = Dmin, col = group), alpha = 1, shape = 22)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none", plot.title = element_text(hjust = 0.5))

Shigh_plot
```


High limits plots
```{r}
#plot_grid(Dhigh_plot, Shigh_plot, Phigh_plot, ncol = 3)
```


Breadth
Development breadth
```{r}

DB_i <- lmer(Dbreadth ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(BreadthLimits)) #
DB_a <- lmer(Dbreadth ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(BreadthLimits)) #

anova(DB_i, DB_a, test.statistic = "Chisq") # keep interaction model, plot interaction
```

```{r}
DB_a_ML <- lmer(Dbreadth ~ -1 + lifestage + absLat +  (1|sp), REML = F, data = as.data.frame(BreadthLimits)) #
summary(DB_a_ML)
#plot(DB_a_ML)
```

```{r}
# Predictions plot
BreadthLimits$group <- factor(BreadthLimits$lifestage)
DbreadthPred <- ggpredict(DB_i, terms = c("absLat","lifestage"))
Dbreadth_plot <- ggplot(DbreadthPred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line(size = 2)+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1, linetype = 2)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(D[Breadth]))+
  xlab("Absolute latitude")+
  ylim(5, 30)+
  annotate("text", x = 0, y = 30, label = "G", size = 5)+
  #geom_point(data = HighLimits, mapping = aes(x = absLat, y = Ph, col = group), alpha = 0.6, shape = 16)+
  #geom_point(data = LowLimits, mapping = aes(x = absLat, y = Dmin, col = group), alpha = 1, shape = 22)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")

Dbreadth_plot
```




Performance breadth
```{r}

PB_i <- lmer(Pbreadth ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(BreadthLimits)) #
PB_a <- lmer(Pbreadth ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(BreadthLimits)) #

anova(PB_i, PB_a, test.statistic = "Chisq") # 
```

```{r}
PB_a_ML <- lmer(Pbreadth ~ -1 + lifestage + absLat +  (1|sp), REML = F, data = as.data.frame(BreadthLimits)) #
summary(PB_a_ML)
#plot(PB_a_ML)

```

```{r}
# Predictions plot
BreadthLimits$group <- factor(BreadthLimits$lifestage)
PbreadthPred <- ggpredict(PB_i, terms = c("absLat","lifestage"))
Pbreadth_plot <- ggplot(PbreadthPred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line(size = 2)+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1, linetype = 2)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(P[Breadth]))+
  xlab("Absolute latitude")+
  ylim(5, 30)+
  annotate("text", x = 0, y = 30, label = "I", size = 5)+
  #geom_point(data = HighLimits, mapping = aes(x = absLat, y = Ph, col = group), alpha = 0.6, shape = 16)+
  #geom_point(data = LowLimits, mapping = aes(x = absLat, y = Dmin, col = group), alpha = 1, shape = 22)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = c(0.2,0.75))

Pbreadth_plot
```

Survival breadth


```{r}

SB_i <- lmer(Sbreadth ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(BreadthLimits)) #
SB_a <- lmer(Sbreadth ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(BreadthLimits)) #

anova(SB_i, SB_a, test.statistic = "Chisq") # 
```

```{r}
SB_i_ML <- lmer(Sbreadth ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), REML = F, data = as.data.frame(BreadthLimits)) #
summary(SB_i_ML)
#plot(SB_i_ML)

```

```{r}
# Predictions plot
BreadthLimits$group <- factor(BreadthLimits$lifestage)
SbreadthPred <- ggpredict(SB_i, terms = c("absLat","lifestage"))
Sbreadth_plot <- ggplot(SbreadthPred, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line(size = 2)+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1, linetype = 2)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(S[Breadth]))+
  xlab("Absolute latitude")+
  ylim(5, 30)+
  annotate("text", x = 0, y = 30, label = "H", size = 5)+
  #geom_point(data = HighLimits, mapping = aes(x = absLat, y = Ph, col = group), alpha = 0.6, shape = 16)+
  #geom_point(data = LowLimits, mapping = aes(x = absLat, y = Dmin, col = group), alpha = 1, shape = 22)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")



Sbreadth_plot
```


Breadth plots
```{r}
#plot_grid(Dbreadth_plot, Sbreadth_plot, Pbreadth_plot, ncol = 3)
```


Run this code in regular script file to make figure:
```{r}
# plot_grid(Dhigh_plot, Shigh_plot, Phigh_plot,
#           Dlow_plot, Slow_plot, Plow_plot,
#           Dbreadth_plot, Sbreadth_plot, Pbreadth_plot, ncol = 3)
```


