---
title: "Calculate development rate variables"
output: html_notebook
---
## Contents
1. Import data
2. Set selection
3. Consolidation of redundant sets
4. Scaling
5. Linear interpolation
6. Obtain Dl, Dh, Dbreadth


```{r}
# load packages
library(tidyverse) # for data wrangling
library(readxl) # to import excel file
library(cowplot)# to format plots
library(psych) # to get geometric mean

```


## 1. Import data 
```{r}

Data <- read_xlsx("~/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", ""))
```


```{r}
# Discard parasitoids and change character to factor 
#unique(Data$status)
Ana <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana <- Ana %>%
  mutate_if(is.character, factor)

```


## 2. Set selection 
To be included in development rate analyses, sets should have at least 4 temperature treatments reporting development time, and form a complete curve (rise, peak, fall). We restricted analyses to four life stages: egg, larva, pupa, and egg to adult.

```{r}
# Create "is_complete", a function to determine whether a set is:
# "Complete" (has rise, peak, fall)
# "Incomplete" (it's a partial curve)
# "No data" (no development rate data available)
Ana$set2 <- Ana$set
is_complete <- function(dat){
  dat <- dat[!is.na(dat$dr) & !is.na(dat$temp),] # Remove treatments with no dr data
 if(nrow(dat) > 3 ){
   y_max <- max(dat$dr, na.rm = T) # get maximum value for s_dr
  x_max <- dat[dat$dr == y_max, "temp"][[1]][1]# get temperature that maximizes dr 
  species <- unique(dat$sp)
  lifestage <- unique(dat$lifestage)
  seti <- unique(dat$set2)
  print(paste("set:",seti))
  #print(species)
  #print(lifestage)
  #print(x_max)
  #print(y_max)
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  # if(sum(cold$dr >=0) && sum(hot$dr >=0)){
  #     print("Complete")
  #   } else {(print("Incomplete"))}
  result = nrow(cold[!is.na(cold$dr),]) && nrow(hot[!is.na(hot$dr),])
  if(result){
    print("Complete")
  } else {
    print("Incomplete")
  }
 } else{
   print("Less than 4 data points")
 }
}

```


```{r, message=FALSE}
# Apply is_complete() to all sets
SetAssessment <- Ana %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(curve = map(data,is_complete)) %>% 
  select(set, curve) %>% 
  unnest(cols = curve) 
unique(SetAssessment$curve)
```


From a total of `r length(unique(Ana$set))`, only  `r nrow(filter(SetAssessment, curve == "Complete"))` had a complete curve for development rate. There were `r nrow(filter(SetAssessment, curve == "Incomplete"))` incomplete sets and `r nrow(filter(SetAssessment, curve == "Less than 4 data points"))` with less than four data points.

```{r}
# Select sets of life stages of interest: egg, larva, pupa, egg to emergence
complete_sets <- SetAssessment %>% 
  filter(curve == "Complete") 
list_complete <- (unique(complete_sets$set))
life_stages <- c("egg", "larva", "pupa", "eggtoemergence")
Ana_dr <- Ana %>% 
  filter(set %in% list_complete) %>% 
  filter(lifestage %in% life_stages)
egg_dr <- Ana_dr %>%
  filter(lifestage == "egg")
larva_dr <- Ana_dr %>%
  filter(lifestage == "larva")
pupa_dr <- Ana_dr %>%
  filter(lifestage == "pupa")
eggtoemergence_dr <- Ana_dr %>%
  filter(lifestage == "eggtoemergence")

Ana_dr$lifestage <- factor(Ana_dr$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
```
There are `r length(unique(Ana_dr$set))` sets of the relevant lifestages (`r length(unique(Ana_dr$sp))` species). Of those, `r length(unique(egg_dr$set))` correspond to egg (N = `r length(unique(egg_dr$sp))` species), `r length(unique(larva_dr$set))` to larva (N = `r length(unique(larva_dr$sp))` species), `r length(unique(pupa_dr$set))`to pupa (N = `r length(unique(pupa_dr$sp))` species) and `r length(unique(eggtoemergence_dr$set))` to eggtoadult (N = `r length(unique(eggtoemergence_dr$sp))`)

## 3. Consolidation of redundant sets
Redundant sets occur because researchers varied experimental conditions such as host plant, humidity and day length, or because data were reported for males and females separately. We considered multiple measurements reported in the same publication that were obtained at the same temperature, of the same ontogenetic stage of individuals of the same population (sp and locality) to be redundant, and we obtained the geometric mean of all treatments (host, sex, day length, humidity,etc) at the same temperature for analyses. 
Consolidated sets kept the number (unique set identifier) of the first set they include.


```{r}
# make a new column that unifies redundant sets
Ana_dr <- Ana_dr %>% 
  mutate(redundancy_id = as.numeric(factor(paste(sp, locality, lifestage, id))))


# Function to get the geometric mean of development rate (dr_c), and mean (dr2)
mean_redundant <- function(dat){
newset <- dat$set [[1]]
dat %>%
  select(temp, sp, dr) %>% 
  group_by(temp) %>% 
  summarise(dr_c = geometric.mean(dr),
            dr2 = mean(dr), 
            newset = newset,
            n_sets = length(dr),.groups = "drop") 
}

# Apply function to Ana_dr

Ana_dr_consolidated <- Ana_dr %>% 
  group_by(redundancy_id) %>% 
  nest() %>% 
  mutate(consolidated = map(data, mean_redundant)) %>% 
  select(redundancy_id, consolidated) %>% 
  unnest(col = consolidated)# Function

# re-name variables. Note that consolidated sets summarize multiple sets and got assigned the number (set id value) of the first set.
names(Ana_dr_consolidated) <- c("redundancy_id", "temp", "dr", "dr2", "set", "n_sets")
```

## 4. Scaling
Create a function to scale development rate
```{r}
scale.dr <- function(table){
  table <- table[!is.na(table$dr) & !is.na(table$temp),] # Remove treatments with no performance data
  if(nrow(table) > 1 ){
  maxdr <- max(table$dr, na.rm = T)
  table$s_dr <- table$dr/maxdr
  table
  }
}


```


Apply the function to each set to obtain s_dr
```{r}
Ana_dr_consolidated <- Ana_dr_consolidated %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(s_dr = map(data, scale.dr)) %>% 
  unnest(cols = s_dr) %>% 
  select(-data)

# ggplot(Ana_dr_consolidated, aes(x = dr, y = s_dr))+
#   geom_point()
```

## 5. Linear interpolation
Obtain predicted development rate in the range from 0 to 40C using linear interpolation. 

```{r}
# 1- Make interpolation function:
interpolate <- function(dat) {
  out <- seq(from = 0, to = 40, by = 0.1)
  as.data.frame(approx(x = dat$temp, y = dat$s_dr, xout = out, method = "linear", rule = 1))
  
}
# 2. Apply interpolation function to all sets
Predicted_Ana_dr <-  Ana_dr_consolidated %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(predicted = map(data,interpolate)) %>% 
  select(set, predicted) %>% 
  unnest(cols = predicted)
names(Predicted_Ana_dr) <- c("set", "temp", "s_dr")
```



## 6. Obtain Dl, Dh, Dbreadth
```{r}

# Function to obtain Dbreadth, Dl and Dh
get_interval <- function(dat){
  dat <- dat[!is.na(dat$s_dr) & !is.na(dat$temp),] # Remove treatments with no s_dr data
  Dmin <- min(dat$temp)
  Dmax <- max(dat$temp)
  y_max <- max(dat$s_dr) # get maximum value for s_dr
  x_max <- dat[dat$s_dr == y_max, "temp"][[1]][[1]]# get temperature that maximizes s_dr
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  limit <- y_max/2
  if (min(cold$s_dr) > limit){
    print("cold incomplete")
    y_mincold <- NA
    x_min_cold <- NA
    grade_cold <- "incomplete"
  } else{
    y_mincold <- min(filter(cold, s_dr >= y_max/2)[["s_dr"]])
    x_min_cold <- min(filter(cold, s_dr >= y_max/2)[["temp"]])
    grade_cold <- "complete"
  }
  
  if(min(hot$s_dr) > limit){
    print("heat incomplete")
    y_minhot <- NA
    x_min_hot <- NA 
    grade_hot <- "incomplete"
  } else {
    y_minhot <- min(filter(hot, s_dr >= y_max/2)[["s_dr"]])
    x_min_hot <- max(filter(hot, s_dr >= y_max/2)[["temp"]])
    grade_hot <- "complete"
  }
  
  
  output <- tibble(Dl = x_min_cold, 
                   Dh = x_min_hot, 
                   Dbreadth = Dh - Dl, 
                   gradecold = grade_cold, 
                   gradehot = grade_hot)
}
```


```{r}
# Apply get_interval() to each set

Interval <- Predicted_Ana_dr %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(interval = map(data,get_interval)) %>% 
  select(set, interval) %>% 
  unnest(cols = interval)

add_info <- Ana_dr %>% 
  select(set, sp, family, lifestage, lat, lon, locality, quality) %>% 
  distinct()
Interval_table1 <- inner_join(Interval, add_info, by = "set")
Interval_table1$lifestage <- factor(Interval_table1$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))

head(Interval_table1)
```

Print table for analysis

```{r}
# write_csv(Interval_table1,"~/Desktop/Projects/ThermalTManuscriptScripts/DevelopmentTable.csv")
```





