---
title: "Hypotheses plots"
output: html_notebook
---



```{r}
remotes::install_github("padpadpadpad/rTPC")
library(rTPC)
library(nls.multstart)
```



```{r, message= FALSE, warning=FALSE}
# load packages
library(tidyverse)
library(readxl)
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
library(devRate)
library(minpack.lm)
```


# Version IV

```{r}
# Import Toy data (based on set 944)

ToyData <- read_xlsx("~/Desktop/Projects/ThermalStressCompilation/Toydata.xlsx", 
                  sheet = "toy2", na = c("NA", ""))
Data <- read_xlsx("~/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", ""))
treats <- Data %>% 
  group_by(set) %>% 
  summarise(count= n())


ggplot(filter(Data, set == 60), aes(x = temp, y = 1/dt))+
  geom_point()
```


```{r}


ex <- Data %>% 
  filter(set == 60) %>% 
  mutate(dr = (1/dt)/0.427) %>% 
  filter(!is.na(dr)) %>% 
  select(set, temp, dr)

d <- ToyData %>% 
  select(temp, dr5) %>% 
  filter( ! is.na(dr5))
a = 
mod = "lactin2_1995"
# get start values
#start_vals <- get_start_vals(d$temp, d$dr5, model_name = "lactin2_1995")
start_vals <- get_start_vals(ex$temp, ex$dr, model_name = "lactin2_1995")
# get limits
#low_lims <- get_lower_lims(d$temp, d$dr5, model_name = "lactin2_1995")
low_lims <- get_lower_lims(ex$temp, ex$dr, model_name = "lactin2_1995")
#upper_lims <- get_upper_lims(d$temp, d$dr5, model_name = "lactin2_1995")
upper_lims <- get_upper_lims(ex$temp, ex$dr, model_name = "lactin2_1995")



fit <- nls_multstart(dr~lactin2_1995(temp = temp, a,b,tmax,delta_t),
                                                     data = ex,
                                                     iter = 500,
                                                     start_lower = start_vals - 10,
                                                     start_upper = start_vals + 10,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y', 
                     convergence_count = FALSE)
fit
linearPortion <- ex[3:9, ]
fitLinear <- lm(dr~ temp, data = linearPortion)

## decorate lm object with a new class lm_left
lm_left <- function(formula,data,...){
  mod <- lm(formula,data)
  class(mod) <- c('lm_left',class(mod))
  mod
}

predictdf.lm_left <- 
  function(model, xseq, se, level){
    init_range = range(model$model$x)
    ## here the main code: truncate to x values at the left
    xseq <- xseq[xseq <=init_range[2]]
    ggplot2:::predictdf.default(model, xseq[-length(xseq)], se, level)
  }
```

```{r}
max(ex$dr)
# predict new data
new_data <- data.frame(temp = seq(5, 40, 0.5))
preds <- augment(fit, newdata = new_data)

# plot data and model fit
 ggplot(new_data, aes(temp, .fitted)) +
  geom_line(aes(temp, .fitted), preds, col = 'grey40', size = 1.5, alpha = 0.5) +
  geom_point(data = ex, mapping = aes(x =temp, y = dr), col = 'black', size = 2, shape = 1) +
  stat_smooth(mapping = aes(temp, dr), data = linearPortion, method = "lm_left", fullrange =T, se = F, col = "black", linetype = 2, size = 2, alpha = 0.5)+
  theme_minimal()+
  ylab("Development rate")+
  xlab("Temperature")+
  ylim(0, 1)+
  xlim(0, 38)+
  #geom_segment(x = 0, y = 0, xend = Inf, yend = 0, size = 1, linetype = 1)+
   
  #annotate("text", x = 9, y = -0.08, label = expression("T"[0]))+
  theme(axis.text = element_text(size = 16))+
  #theme(axis.text.x= c("Tmin", "T0", "Topt", "Tmax"))+
  #scale_y_continuous(expand = expansion(mult = c(0, 0)))+
 scale_x_continuous(breaks = c(11, 13, 32, 36), labels = c(expression("T"[min]), expression("T"[0]), expression("T"[opt]), expression("T"[max])), limits = c(5, 40))
   
  #theme(axis.ticks.x = element_blank(), axis.title.y = element_text(colour = "grey40"), axis.title.y.left = element_text(colour = "black"))+
  #scale_y_continuous(name = "Development rate", sec.axis = sec_axis(~ .*1, name = "Survival"),breaks = c(0, 0.5, 1))
  #scale_y_continuous( breaks = c(0, 1))
  

```


When the interval of temperatures that allow for 50% survival is larger on both the cold and the hot side than the interval that allows for development at 50% efficiency, performance is entirely limited by development rate.

```{r}

Figure1

```


Row 1

```{r}
ToyData$p1 <- ToyData$dr2 *(ToyData$shigher/100)
ToyData$p2 <- ToyData$dr2 *(ToyData$highequal/100)
ToyData$p3 <- ToyData$dr3 *(ToyData$slower/100)
ToyData$p4 <- ToyData$dr4 *(ToyData$slower/100)
ToyData$p5 <- ToyData$dr5 *(ToyData$slower/100)
A1 <- ggplot(ToyData, aes(x = temp, y = p1))+
  # performance
  geom_line(data = ToyData, mapping = aes(x = temp, y = p1), size = 2, col = "#90548bff", alpha = 1)+ # performance, orange:"#E55C30FF", green: "#5DC863FF"
  #geom_segment(x = 17.5, y = 0.5, xend = 31.3, yend = 0.5, linetype = 1, col =  "#90548bff", size = 2, alpha = 1)+
  annotate("rect", xmin = 17.5, xmax = 31.3, ymin = -Inf, ymax = 1, fill =  "#90548bff", alpha = 0.3)+
  
  # survival
   geom_line(data = ToyData, mapping = aes(x = temp, y = shigher/100), size = 1, linetype = 1, col = "grey40")+ # survival
   geom_segment(x = 9.1, y = 0.49, xend = 34.8, yend = 0.49, linetype = 1, col = "grey40")+
  
  # development
  geom_line(data = ToyData, mapping = aes(x = temp, y = dr2), size = 2, col = "black", linetype = 3)+ #development rate
  geom_segment(x = 17.5, y = 0.51, xend = 31.3, yend = 0.51, linetype = 3)+
  
  
  # Points
  geom_point(x = 9.1, y = 0.5, size = 4, shape = 19, col = "blue")+
  geom_point(x = 34.8, y = 0.5, size = 4, shape = 15, col = "orange")+
  geom_point(x = 17.5, y = 0.5, size = 6, shape = 1, col = "blue", alpha = 0.5)+
  geom_point(x = 31.3, y = 0.5, size = 6, shape = 0, col = "orange", alpha = 0.5)+
 
  
  ylab("Rate")+
  xlab("")+
 
  annotate("text", x = 17.5, y = 0.01, label = expression("P"[l]), size = 5) +
  annotate("text", x = 31.3, y = 0.01, label = expression("P"[h]), size = 5) +
  
   scale_x_continuous(breaks = c(9.1, 17.5,  31.3, 34.8), labels = c(expression("S"[l]), expression("D"[l]), expression("D"[h]), expression("S"[h])))+
   theme(axis.title.y = element_text(colour = "black"), axis.title.y.left = element_text(colour = "black"))+
  scale_y_continuous(name = "Rate",
         #sec.axis = sec_axis(~ .*1, name = "Survival",breaks = c(0, 0.5, 1)),
                     breaks = c(0, 0.5, 1))+
  theme_cowplot()
 

A1
```


A2
```{r}
A1data <- tibble(survival = c(34.8,9.1),
                 Limit = factor(c("high", "low")),
                 development = c(31.3, 17.5)
                 )
A2 <- ggplot(A1data, aes(x = survival, y = development, shape = Limit, col = Limit))+
  geom_point(size = 5, alpha = 0.5)+
  geom_point(size = 3)+
  scale_shape_manual(values = c(15,16))+
  scale_color_manual(values = c("orange", "blue"))+
  ylim(0,40)+
  xlim(0,40)+
  ylab("D limits (°C)")+
  xlab("")+
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "black", linetype = 2, size =0.5, alpha = 0.1)+
  theme_cowplot()+
  #theme(legend.position = c(0.6,0.4))
  theme(legend.position = "none")
A2
```


When development limits (Dlow and Dhigh) are equal to survival limits (Slow and Shigh), performance interval is reduced, with Plow being higher than Dlow and Slow; and Phigh smaller than Dhigh and Shigh

Second row
```{r}
label1 <- expression("D"[low])
label2 <- expression("S"[low])
B1 <- ggplot(ToyData, aes(x = temp, y = p4))+
 
  # performance
  geom_line(size = 2, col = "#90548bff")+ # performance
  #geom_segment(x = 23.7, y = 0.5, xend = 32.8, yend = 0.5, linetype = 1, col = "#90548bff", size = 2)+
 # annotate("text", x = 24, y = 0.5, label = expression("P"[low]))+
  #annotate("text", x = 33, y = 0.5, label = expression("P"[high]))+
  annotate("rect", xmin = 23.7, xmax = 32.8, ymin = -Inf, ymax = 1, fill = "#90548bff", alpha = 0.3)+
  
  
  # survival
  geom_line(data = ToyData, mapping = aes(x = temp, y = lowequal/100), size = 1, linetype = 1, col = "grey40")+ # survival

  # development
  geom_line(data = ToyData, mapping = aes(x = temp, y = dr3), size = 2, col = "black", linetype = 3)+ #development rate
  #geom_segment(x = 20.5, y = 0.51, xend = 34.2, yend = 0.51, linetype = 3)+
  #geom_segment(x = 20.5, y = -0.16, xend = 34.2, yend = -0.16, linetype = 3)+
  #annotate("text", x = 19., y = -0.05, label = expression("D"[low]))+
  #annotate("text", x = 35.5, y = -0.05, label = expression("D"[high]))+
  
  
  geom_segment(x = 20.5, y = 0.51, xend = 34.2, yend = 0.51, linetype = 3, col = "black")+
  geom_segment(x = 20.5, y = 0.49, xend = 34.2, yend = 0.49, linetype = 1, col = "grey40")+
 # geom_segment(x = 20.5, y = 0, xend = 34.2, yend = 0, linetype = 1, col = "grey40")+
  annotate("text", x = 20.5, y = 0, label = expression("S"[l]), size = 4)+
  annotate("text", x = 34.2, y = 0, label = expression("S"[h]), size = 4)+
  
  annotate("text", x = 23.7, y = 0.1, label = expression("P"[l]), size = 5) +
  annotate("text", x = 32.8, y = 0.1, label = expression("P"[h]), size = 5) +
  #geom_segment(x = 9.5, y = 0.49, xend = 39, yend = 0.49, linetype = 2)+
  theme_cowplot()+
  #annotate("text", x =5, y = 1, label = "B")+
  
   geom_point(x = 20.5, y = 0.5, size = 6, shape = 1, col = "blue")+
   geom_point(x = 34.2, y = 0.5, size = 6, shape = 0, col = "orange")+
   geom_point(x = 20.5, y = 0.5, size = 4, shape = 19, col = "blue")+
   geom_point(x = 34.2, y = 0.5, size = 4, shape = 15, col = "orange")+
  
  #ylab("D limits (°C)")+
  xlab("")+
  scale_y_continuous(name = "Rate", 
                     #sec.axis = sec_axis(~ .*1, name = "Survival",breaks = c(0, 0.5, 1)),
                     breaks = c(0, 0.5, 1))+
  scale_x_continuous(breaks = c(20.5, 34.2), labels = c(expression("D"[l]), expression("D"[h])))
  #theme(axis.ticks.x = element_blank(),axis.ticks.y.left = element_blank(), axis.title.y.left = element_blank(), axis.title.y.right = element_text(colour = "black"),axis.text.y.left = element_blank() , axis.line.y.left = element_blank())
# scale_x_continuous(breaks = c(9.5, 27, 39.5), labels = c("chill stress", "optimal range", "heat stress"))+
  # theme(axis.ticks.x = element_blank())
B1
```



```{r}
B1data <- tibble(survival = c(34.5,20.5),
                 Limit = factor(c("high", "low")),
                 development = c(34.5, 20.5)
                 )
B2 <- ggplot(B1data, aes(x = survival, y = development, shape = Limit, col = Limit))+
  geom_point(size = 5, alpha = 0.5)+
  geom_point(size = 3)+
  scale_shape_manual(values = c(15,16))+
  scale_color_manual(values = c("orange", "blue"))+
  ylim(0,40)+
  xlim(0,40)+
  ylab("D limits (°C)")+
  xlab("")+
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "black", linetype = 2, size =0.5, alpha = 0.1)+
  theme_cowplot()+
  theme(legend.position = "none")
B2
```

3A
```{r}

C1 <- ggplot(ToyData, aes(x = temp, y = p3))+
 
  
  
  # performance
  geom_line(size = 2, col = "#90548bff")+ # performance
  #geom_segment(x = 20.5, y = 0.5, xend = 31.8, yend = 0.5, linetype = 1, col = "#90548bff", size = 2)+
  annotate("rect", xmin = 20.5, xmax = 31.8, ymin = -Inf, ymax = 1, fill = "#90548bff", alpha = 0.3)+
  annotate("text", x = 20.5, y = 0.01, label = expression("P"[l]), size = 5)+
  annotate("text", x = 31.8, y = 0.01, label = expression("P"[h]), size = 5)+
  
  # survival
  geom_line(data = ToyData, mapping = aes(x = temp, y = slower/100), size = 1, linetype = 1, col = "grey40")+ # survival
  geom_segment(x = 9.1, y = 0.49, xend = 32.3, yend = 0.49, linetype = 1, col = "grey40")+
  #geom_segment(x = 9.1, y = -0.24, xend = 32.3, yend = -0.24, linetype = 1, col = "grey40")+
  #annotate("text", x = 8.5, y = -0, label = expression("S"[low]))+
  #annotate("text", x = 33.8, y = -0, label = expression("S"[high]))+
  #geom_segment(x = 9.5, y = 0.49, xend = 39, yend = 0.49, linetype = 2)+
  
  
  # development
  geom_line(data = ToyData, mapping = aes(x = temp, y = dr3), size = 2, col = "black", linetype = 3)+ 
  
  #development rate
  geom_segment(x = 20.5, y = 0.51, xend = 34.2, yend = 0.51, linetype = 3)+
  #geom_segment(x = 20.5, y = -0.16, xend = 34.2, yend = -0.16, linetype = 3)+
  #annotate("text", x = 19.5, y = 0, label = expression("D"[low]))+
  #annotate("text", x = 36, y = 0, label = expression("D"[high]))+
 
  
  
  
  theme_cowplot()+
  #annotate("text", x =5, y = 1, label = "C")+
  ylab("Development rate")+
  xlab("")+
  geom_point(x = 20.5, y = 0.5, size = 6, shape = 1, col = "blue")+
   geom_point(x = 34.2, y = 0.5, size = 6, shape = 0, col = "orange")+
   geom_point(x = 9.1, y = 0.5, size = 4, shape = 19, col = "blue")+
   geom_point(x = 32.3, y = 0.5, size = 4, shape = 15, col = "orange")+
  #ylim(-0.25,1)+
  #annotate("rect", xmin = 20.5, xmax = 31.8, ymin = 0, ymax = 1, fill = "#90548bff", alpha = 0.3)+
  # annotate("text", x = 12, y = -0.08, label = "chill stress")+
  # annotate("text", x = 39.5, y = -0.08, label = "heat stress")+
  # annotate("text", x = 28, y = -0.08, label = "optimal range")+
  # 
  #geom_segment(x = 12, y = 0, xend = 29, yend = 1)+
  #annotate("text", x = 11, y = -0.08, label = expression("T"[0]))+
  #theme(axis.text = element_text(size = 16))
  #theme(axis.text.x= c("Tmin", "T0", "Topt", "Tmax"))+
  #scale_x_continuous(breaks = c(8, 11, 31, 36), labels = c(expression("T"[min]), expression("T"[0]), expression("T"[opt]), expression("T"[high])))
  #scale_x_continuous(breaks = c(9.5, 20.5, 39.5), labels = c("chill stress", "optimal range", "heat stress"))
  scale_x_continuous(breaks = c(9.1,20.5,32.3,34.2), labels = c(expression("S"[l]), expression("D"[l]), expression("S"[h]), expression("D"[h])))+
  scale_y_continuous(name = "Rate", 
                     #sec.axis = sec_axis(~ .*1, name = "Survival",breaks = c(0, 0.5, 1)),
                     breaks = c(0, 0.5, 1))
  
C1


```


```{r}
C1data <- tibble(survival = c(32.3,9.1),
                 Limit = factor(c("high", "low")),
                 development = c(34.2, 19.5)
                 )
C2 <- ggplot(C1data, aes(x = survival, y = development, shape = Limit, col = Limit))+
  geom_point(size = 5, alpha = 0.5)+
  geom_point(size = 3)+
  scale_shape_manual(values = c(15,16))+
  scale_color_manual(values = c("orange", "blue"))+
  ylim(0,40)+
  xlim(0,40)+
  ylab("D limits (°C)")+
  xlab("")+
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "black", linetype = 2, size =0.5, alpha = 0.1)+
  theme_cowplot()+
  theme(legend.position = "none")
C2
```



D1
```{r}
D1 <- ggplot(ToyData, aes(x = temp, y = p5))+
  
 # performance
  geom_line(size = 2, col = "#90548bff")+ # performance
  #geom_segment(x = 24.8, y = 0.5, xend = 34.2, yend = 0.5, linetype = 1, col = "#90548bff", size = 2)+
  annotate("rect", xmin = 24.8, xmax = 34.2, ymin = -Inf, ymax = 1, fill = "#90548bff", alpha = 0.3)+
  annotate("text", x = 24.8, y = 0.01, label = expression("P"[l]), size = 5)+
  annotate("text", x = 34.2, y = 0.01, label = expression("P"[h]), size = 5)+
  
  
  # development
  geom_line(data = ToyData, mapping = aes(x = temp, y = dr3), size = 2, col = "black", linetype = 3)+ #development rate
  geom_segment(x = 20.5, y = 0.51, xend = 34.2, yend = 0.51, linetype = 3)+
 # geom_segment(x = 20.5, y = -0.16, xend = 34.2, yend = -0.16, linetype = 3)+
  #annotate("text", x = 19., y = 0, label = expression("D"[low]))+
  #annotate("text", x = 35., y = 0, label = expression("D"[high]))+
  # survival
  geom_line(data = ToyData, mapping = aes(x = temp, y = surhigher/100), size = 1, linetype = 1, col = "grey40")+ # survival
  geom_segment(x = 22.3, y = 0.49, xend = 37, yend = 0.49, linetype = 1, col = "grey40")+
  #geom_segment(x = 22.3, y = -0.24, xend = 37, yend = -0.24, linetype = 1, col = "grey40")+
 # annotate("text", x = 21.5, y = 0, label = expression("S"[low]))+
  #annotate("text", x = 38, y = 0, label = expression("S"[high]))+
  
    
    #geom_segment(x = 9.5, y = 0.49, xend = 39, yend = 0.49, linetype = 2)+
  theme_cowplot()+
 # annotate("text", x =5, y = 1, label = "D")+
  ylab("Rate")+
  xlab("Temperature (°C)")+
  #ylim(-0.25,1)+
  geom_point(x = 20.5, y = 0.5, size = 6, shape = 1, col = "blue")+
   geom_point(x = 34.2, y = 0.5, size = 6, shape = 0, col = "orange")+
   geom_point(x = 22.3, y = 0.5, size = 4, shape = 19, col = "blue")+
   geom_point(x = 37, y = 0.5, size = 4, shape = 15, col = "orange")+
  # annotate("text", x = 12, y = -0.08, label = "chill stress")+
  # annotate("text", x = 39.5, y = -0.08, label = "heat stress")+
  # annotate("text", x = 28, y = -0.08, label = "optimal range")+
  # 
  #geom_segment(x = 12, y = 0, xend = 29, yend = 1)+
  #annotate("text", x = 12, y = -0.08, label = expression("T"[0]))+
  #theme(axis.text = element_text(size = 16))
  #theme(axis.text.x= c("Tmin", "T0", "Topt", "Tmax"))+
  scale_y_continuous(  
    #sec.axis = sec_axis(~ .*1, name = "Survival", breaks = c(0, 0.5, 1)),
    breaks = c(0, 0.5, 1))+
  scale_x_continuous(breaks = c(20.5, 22.3, 34.2, 37), labels = c(expression("D"[l]),expression("S"[l]),expression("D"[h]), expression("S"[h])))
  #theme(axis.ticks.x = element_blank(),axis.ticks.y.left = element_blank(), axis.title.y.left = element_blank(), axis.title.y.right = element_text(colour = "black"),axis.text.y.left = element_blank(), axis.line.y.left = element_blank())
# scale_x_continuous(breaks = c(9.5, 27, 39.5), labels = c("chill stress", "optimal range", "heat stress"))+
  # theme(axis.ticks.x = element_blank())

D1

```



```{r}
D1data <- tibble(survival = c(37,22.3),
                 Limit = factor(c("high", "low")),
                 development = c(34.2, 20.5)
                 )
D2 <- ggplot(D1data, aes(x = survival, y = development, shape = Limit, col = Limit))+
  geom_point(size = 5, alpha = 0.5)+
  geom_point(size = 3)+
  scale_shape_manual(values = c(15,16))+
  scale_color_manual(values = c("orange", "blue"))+
  ylim(0,40)+
  xlim(0,40)+
  ylab("D limits (°C)")+
  xlab("S limits (°C)")+
  
  geom_segment(x = 0, xend = 40, y = 0, yend = 40, col = "black", linetype = 2, size =0.5, alpha = 0.1)+
  theme_cowplot()+
  theme(legend.position = "none")
D2
```

Figure2
```{r}
F2 <- plot_grid(A1, A2, 
                  B1, B2,
                  C1,  C2,
                  D1,  D2, ncol = 2, rel_widths = c(
                  1, 0.75,
                  1, 0.75,
                  1, 0.75,
                  1, 0.75), labels= "AUTO")

F2
```






```{r}
Figure_1Updated <- ggplot(ToyData, aes(x = temp, y = dr4))+
  
 # performance
  # geom_line(size = 2, col = "#90548bff", alpha = 1)+ # performance, orange:"#E55C30FF", green: "#5DC863FF"
  # geom_segment(x = 17.5, y = -0.08, xend = 31.3, yend = -0.08, linetype = 1, col =  "#90548bff", size = 2, alpha = 1)+
  # annotate("rect", xmin = 17.5, xmax = 31.3, ymin = 0, ymax = 1, fill =  "#90548bff", alpha = 0.3)+
  # annotate("text", x = 16, y = -0.08, label = expression("P"[low]))+
  # annotate("text", x = 33, y = -0.08, label = expression("P"[high]))+
  # 
  # development
  #geom_line(data = ToyData, mapping = aes(x = temp, y = dr5), size = 2, col = "darkgrey", linetype = 1)+
  geom_smooth(data = ToyData, mapping = aes(x = temp, y = p5), se = F, size = 2, col = "darkgrey", linetype = 1)+
  #geom_segment(x = 17.5, y = 0.51, xend = 31.3, yend = 0.51, linetype = 3)+
  #geom_segment(x = 17.5, y = -0.16, xend = 31.3, yend = -0.16, linetype = 3)+
  #annotate("text", x = 16, y = -0.16, label = expression("D"[low]))+
  #annotate("text", x = 33, y = -0.16, label = expression("D"[high]))+
 
  # # survival
  #  geom_line(data = ToyData, mapping = aes(x = temp, y = shigher/100), size = 1, linetype = 1, col = "grey40")+ # survival
  #  geom_segment(x = 9.1, y = 0.49, xend = 34.8, yend = 0.49, linetype = 1, col = "grey40")+
  # geom_segment(x = 9.1, y = -0.24, xend = 34.8, yend = -0.24, linetype = 1, col = "grey40")+
  # annotate("text", x = 8, y = -0.24, label = expression("S"[low]))+
  # annotate("text", x = 36, y = -0.24, label = expression("S"[high]))+
  #geom_segment(x = 9.5, y = 0.49, xend = 39, yend = 0.49, linetype = 2)+
  theme_cowplot()+
  #annotate("text", x =5, y = 1, label = "A")+
  ylab("Development rate")+
  xlab("Temperature")+
 
 
  
  # annotate("text", x = 12, y = -0.08, label = "chill stress")+
  # annotate("text", x = 39.5, y = -0.08, label = "heat stress")+
  # annotate("text", x = 28, y = -0.08, label = "optimal range")+
  
  # T0
  geom_segment(x = 9, y = 0.0, xend = 26, yend = 1)+
  #annotate("text", x = 9, y = -0.08, label = expression("T"[0]))+
  theme(axis.text = element_text(size = 16))+
  #theme(axis.text.x= c("Tmin", "T0", "Topt", "Tmax"))+
  scale_x_continuous(breaks = c(5, 9, 26, 32), labels = c(expression("T"[min]), expression("T"[0]), expression("T"[opt]), expression("T"[max])))+
  #scale_x_continuous(breaks = c(9.5, 20.5, 39.5), labels = c("chill stress", "optimal range", "heat stress"))
   
  # theme(axis.ticks.x = element_blank(), axis.title.y = element_text(colour = "grey40"), axis.title.y.left = element_text(colour = "black"))+
  #scale_y_continuous(name = "Development rate", sec.axis = sec_axis(~ .*1, name = "Survival"),breaks = c(0, 0.5, 1))
  scale_y_continuous(breaks = c(0, 0.5, 1))
 
#theme(axis.title.x = element_text(colour = "red"))



```



