---
title: "Calculation of Performance variables"
output: html_notebook
---

We used r (tydiverse) for data wrangling, (ggplot2) for data visualization, (psych) for obtaining geometric mean, (readxl) to import data files and (cowplot) to format ggplots.

```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
library(psych)
```

Import data 
```{r}
# Mariana's desktop
Data <- read_xlsx("~/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", ""))
```


```{r}
# Discard parasitoids and change character to factor 
unique(Data$status)
Ana_1 <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana_1 <- Ana_1 %>%
  mutate_if(is.character, factor)
```

# 2. Calculate P (only for sets with both survival and development time)

2.1 Rescale D and S
```{r}
scale.dr <- function(table){
  table <- table[!is.na(table$dr) & !is.na(table$temp),] # Remove treatments with no performance data
  if(nrow(table) > 1 ){
  maxdr <- max(table$dr, na.rm = T)
  table$s_dr <- table$dr/maxdr
  table
  }
}

scale.su <- function(table){
  table <- table[!is.na(table$survival) & !is.na(table$temp),] # Remove treatments with no performance data
  if(nrow(table) > 1 ){
  maxsur <- max(table$survival, na.rm = T)[[1]]
  table$s_sur <- table$survival/maxsur
  table
  }
}

```


Apply the function to each set and add s_dr and s_sur
```{r}
Ana_dr <- Ana_1 %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(s_dr = map(data, scale.dr)) %>% 
  unnest(cols = s_dr) %>% 
  select(-data)

Ana_sur <- Ana_1 %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(s_sur = map(data, scale.su)) %>% 
  unnest(cols = s_sur) %>% 
  select(-data) %>% 
  select(set, s_sur, temp)

Ana <- left_join(Ana_dr, Ana_sur, by = c("set", "temp"))
```



Multiply s_dr and s_sur to obtain P
```{r}
Ana$P_survival <- Ana$s_dr * (Ana$survival/100)
Ana$P <- Ana$s_dr * Ana$s_sur

# plots for sanity check
#ggplot(Ana, aes(x = dr, y = s_dr))+
#  geom_point()

#ggplot(Ana, aes(x = survival, y = s_sur))+
#  geom_point()

```


# 3. Select sets with at least 4 data points and a complete curve (rise, peak, fall) for egg, larva, pupa, and egg to adult.

To be included in performance analyses, sets should have at least 4 temperature treatments and form a complete P curve (rise, peak, fall). We restricted analyses to four life stages: egg, larva, pupa and egg to adult.

```{r}
# Function to determine whether a set is:
# "Complete" (has low, peak and high values)
# "Incomplete" (only low and peak)
# "No data" (no P data available)
Ana$set2 <- Ana$set
is_complete <- function(dat){
  dat <- dat[!is.na(dat$P) & !is.na(dat$temp),] # Remove treatments with no performance data
 if(nrow(dat) > 3 ){
  
   y_max <- max(dat$P, na.rm = T) # get maximum value for performance
  x_max <- dat[dat$P == y_max, "temp"][[1]][1]# get temperature that maximizes performance 
  
  species <- unique(dat$sp)
  lifestage <- unique(dat$lifestage)
  seti <- unique(dat$set2)
  print(paste("set:",seti))
  #print(species)
  #print(lifestage)
  #print(x_max)
  #print(y_max)
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  result = nrow(cold[!is.na(cold$P),]) && nrow(hot[!is.na(hot$P),])
  if(result){
    print("Complete")
  } else {
    print("Incomplete")
  }
 } else{
   print("Less than 4 data points")
 }
  
  
}

```


```{r}
# Apply is_complete() to all sets
SetAssessment <- Ana %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(curve = map(data,is_complete)) %>% 
  select(set, curve) %>% 
  unnest(cols = curve) 

```


From a total of `r length(unique(Ana$set))`, only  `r nrow(filter(SetAssessment, curve == "Complete"))` had a complete curve for performance. There were `r nrow(filter(SetAssessment, curve == "Incomplete"))` incomplete sets and `r nrow(filter(SetAssessment, curve == "Less than 4 data points"))` with less than four data points
```{r}
# Select sets of life stages of interest: egg, larva, pupa, egg to emergence
complete_sets <- SetAssessment %>% 
  filter(curve == "Complete") 
list_complete <- (unique(complete_sets$set))
life_stages <- c("egg", "larva", "pupa", "eggtoemergence")
Ana_performance <- Ana %>% 
  filter(set %in% list_complete) %>% 
  filter(lifestage %in% life_stages)
egg_performance <- Ana_performance %>%
  filter(lifestage == "egg")
larva_performance <- Ana_performance %>%
  filter(lifestage == "larva")
pupa_performance <- Ana_performance %>%
  filter(lifestage == "pupa")
eggtoemergence_performance <- Ana_performance %>%
  filter(lifestage == "eggtoemergence")

Ana_performance$lifestage <- factor(Ana_performance$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
```

There are `r length(unique(Ana_performance$set))` sets of the relevant lifestages (`r length(unique(Ana_performance$sp))` species). Of those, `r length(unique(egg_performance$set))` correspond to egg (N = `r length(unique(egg_performance$sp))` species), `r length(unique(larva_performance$set))` to larva (N = `r length(unique(larva_performance$sp))` species), `r length(unique(pupa_performance$set))`to pupa (N = `r length(unique(pupa_performance$sp))` species) and `r length(unique(eggtoemergence_performance$set))` to eggtoadult (N = `r length(unique(eggtoemergence_performance$sp))`)

#3. Consolidate redundant sets
Redundant sets occur beacuse researchers varied experimental conditions such as host plant, humidity and daylength, or because data were reported for males and females separately. We considered multiple measurements reported in the same publication that were obtained at the same temperature, of the same ontogenetic stage of individuals of the same population (sp and locality) to be redundant, and we obtained the geometric mean of all treatments (host, sex, daylength, humidity,etc) at the same temperature for analyses. 
Consolidated sets kept the number (unique set identifier) of the first set they include.


```{r}
# make a new column that unifies redundant sets
# Ana_performance <- Ana_performance %>% 
#   mutate(redundancy_id = as.numeric(factor(paste(sp, locality, lifestage, id))))
Ana_performance$redundancy_id <- as.numeric(factor(paste(Ana_performance$sp, Ana_performance$locality, Ana_performance$lifestage, Ana_performance$id)))


#length(unique(Ana_performance$set)) # there were 139 sets
#length(unique(Ana_performance$redundancy_id)) # there were 137 non-redundant sets

# Function to get the geometric mean of development rate (dr_c), and mean (dr2)
mean_redundant <- function(dat){
newset <- dat$set [[1]]
dat %>%
  select(temp, sp, P) %>% 
  group_by(temp) %>% 
  summarise(performance_c = geometric.mean(P),
            performance2 = mean(P), 
            newset = newset,
            n_sets = length(P),.groups = "drop") 
}

# Apply function to Ana_performance

Ana_performance_consolidated <- Ana_performance %>% 
  group_by(redundancy_id) %>% 
  nest() %>% 
  mutate(consolidated = map(data, mean_redundant)) %>% 
  select(redundancy_id, consolidated) %>% 
  unnest(col = consolidated)# Function

# re-name variables. Note that consolidated sets summarize multiple sets and got assigned the number (set id value) of the first set.
names(Ana_performance_consolidated) <- c("redundancy_id", "temp", "P", "P2", "set", "n_sets")
```


4. Obtain predicted performance in the range from 0 to 40C using linear interpolation. 

```{r}
#create curves: 1- Make interpolation function:
interpolate <- function(dat) {
  out <- seq(from = 0, to = 40, by = 0.1)
  as.data.frame(approx(x = dat$temp, y = dat$P, xout = out, method = "linear", rule = 1))
  
}

Predicted_Ana_performance <-  Ana_performance_consolidated %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(predicted = map(data,interpolate)) %>% 
  select(set, predicted) %>% 
  unnest(cols = predicted)
names(Predicted_Ana_performance) <- c("set", "temp", "P")
```



#7. Calculate P-interval (min, max, length)


```{r}


get_interval <- function(dat){
  dat <- dat[!is.na(dat$P) & !is.na(dat$temp),] # Remove treatments with no P data
  Pmin <- min(dat$temp)
  Pmax <- max(dat$temp)
  y_max <- max(dat$P) # get maximum value for dr
  x_max <- dat[dat$P == y_max, "temp"][[1]][[1]]# get temperature that maximizes P
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  limit <- y_max/2
  print(limit)
  if (min(cold$P) > limit){
    print("cold incomplete")
    y_mincold <- NA
    x_min_cold <- NA
    grade_cold <- "incomplete"
  } else{
    y_mincold <- min(filter(cold, P >= y_max/2)[["P"]])
    x_min_cold <- min(filter(cold, P >= y_max/2)[["temp"]])
    grade_cold <- "complete"
  }
  
  if(min(hot$P) > limit){
    print("heat incomplete")
    y_minhot <- NA
    x_min_hot <- NA 
    grade_hot <- "incomplete"
  } else {
    y_minhot <- min(filter(hot, P >= y_max/2)[["P"]])
    x_min_hot <- max(filter(hot, P >= y_max/2)[["temp"]])
    grade_hot <- "complete"
  }
  
  
  output <- tibble(Pmin = Pmin,
                   Pmax = Pmax, 
                   Pl = x_min_cold, 
                   Ph = x_min_hot, 
                   Pbreadth = Ph - Pl, 
                   Pgradecold = grade_cold, 
                   Pgradehot = grade_hot)
  
  
}




```


```{r}
# Apply get interval to everything
# apply function to all sets
Interval <- Predicted_Ana_performance %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(interval = map(data,get_interval)) %>% 
  select(set, interval) %>% 
  unnest(cols = interval)
add_info <- Ana_performance %>% 
  select(set, sp, family, lifestage, lat, lon, locality, quality) %>% 
  distinct()
Interval_table1 <- inner_join(Interval, add_info, by = "set")
Interval_table1$lifestage <- factor(Interval_table1$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
# Interval_table_cold <- filter(Interval_table1, gradecold == "complete")
# Interval_table_hot <- filter(Interval_table1, gradehot == "complete")
# Interval_table_both <- filter(Interval_table1, gradecold == "complete", gradehot == "complete")
# rm(Interval_table1)
```



```{r}
 ggplot(Interval_table1, aes(x = abs(lat), y = Pmin))+
  geom_point(col = "blue")+
  geom_point(data = Interval_table1, mapping = aes(x = abs(lat), y = Pl),col = "grey", shape = 21)+
  geom_point(data = Interval_table1, mapping = aes(x = abs(lat), y = Pmax), col = "red", shape = 17)+
  geom_point(data = Interval_table1, mapping = aes(x = abs(lat), y = Ph), col = "orange", shape = 24)



```



```{r}
# write_csv(Performance,"~/Desktop/Projects/ThermalPerformance/PerformanceS.csv")
#write_csv(Interval_table1,"~/Desktop/Projects/ThermalPerformance/PerformanceTable.csv")
```
