---
title: "Calculate survival responses"
output: html_notebook
---
## Contents
1. Import data
2. Set selection
3. Consolidation of redundant sets
4. Linear interpolation
5. Obtain Sl, Sh, Sbreadth

```{r}
# load packages
library(tidyverse)
library(readxl)
library(cowplot)
library(psych)
```

## 1. Import data 
```{r}
Data <- read_xlsx("~/Desktop/Projects/ThermalPerformance/PhysiologyDatabaseVersion5.xlsx", 
                  sheet = "WorkingTable", na = c("NA", ""))
```


```{r}
# Discard parasitoids and change character to factor 
#unique(Data$status)
Ana <- Data[Data$status != "parasitoid",]
rm(Data)

# convert character to factor
Ana <- Ana %>%
  mutate_if(is.character, factor)
```

## 2. Set selection
To be included in survival analyses, sets should have at least 4 temperature treatments reporting survival and form a complete curve (rise, peak, fall). We restricted analyses to four life stages: egg, larva, pupa and egg to adult.

```{r}
# Function to determine whether a set is:
# "Complete" (has low, peak and high values)
# "Incomplete" (only low and peak)
# "No data" (no survival data available)
Ana$set2 <- Ana$set
is_complete <- function(dat){
  dat <- dat[!is.na(dat$survival) & !is.na(dat$temp),] # Remove treatments with no survival data
 if(nrow(dat) > 3 ){
  
   y_max <- max(dat$survival, na.rm = T) # get maximum value for survival
  x_max <- dat[dat$survival == y_max, "temp"][[1]][1]# get temperature that maximizes survival 
  
  species <- unique(dat$sp)
  lifestage <- unique(dat$lifestage)
  seti <- unique(dat$set2)
  print(paste("set:",seti))
  #print(species)
  #print(lifestage)
  #print(x_max)
  #print(y_max)
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  # if(sum(cold$survival >=0) && sum(hot$survival >=0)){
  #     print("Complete")
  #   } else {(print("Incomplete"))}
  result = nrow(cold[!is.na(cold$survival),]) && nrow(hot[!is.na(hot$survival),])
  if(result){
    print("Complete")
  } else {
    print("Incomplete")
  }
 } else{
   print("Less than 4 data points")
 }
  
  
}


```


```{r}
# Apply is_complete() to each set
SetAssessment <- Ana %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(curve = map(data,is_complete)) %>% 
  select(set, curve) %>% 
  unnest() 

```

From a total of `r length(unique(Ana$set))`, only  `r nrow(filter(SetAssessment, curve == "Complete"))` had a complete curve for survival. There were `r nrow(filter(SetAssessment, curve == "Incomplete"))` incomplete sets and `r nrow(filter(SetAssessment, curve == "Less than 4 data points"))` with less than four data points.

```{r}
# Select sets of life stages of interest: egg, larva, pupa, egg to emergence
complete_sets <- SetAssessment %>% 
  filter(curve == "Complete") 
list_complete <- (unique(complete_sets$set))
life_stages <- c("egg", "larva", "pupa", "eggtoemergence")
Ana_survival <- Ana %>% 
  filter(set %in% list_complete) %>%  # this is to get only complete sets
  filter(lifestage %in% life_stages) # and only the life stages of interest
egg_survival <- Ana_survival %>%
  filter(lifestage == "egg")
larva_survival <- Ana_survival %>%
  filter(lifestage == "larva")
pupa_survival <- Ana_survival %>%
  filter(lifestage == "pupa")
eggtoemergence_survival <- Ana_survival %>%
  filter(lifestage == "eggtoemergence")

Ana_survival$lifestage <- factor(Ana_survival$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
```
There are `r length(unique(Ana_survival$set))` sets of the relevant lifestages (`r length(unique(Ana_survival$sp))` species). Of those, `r length(unique(egg_survival$set))` correspond to egg (N = `r length(unique(egg_survival$sp))` species), `r length(unique(larva_survival$set))` to larva (N = `r length(unique(larva_survival$sp))` species), `r length(unique(pupa_survival$set))`to pupa (N = `r length(unique(pupa_survival$sp))` species) and `r length(unique(eggtoemergence_survival$set))` to eggtoadult (N = `r length(unique(eggtoemergence_survival$sp))`)

## 3. Consolidation of redundant sets
Redundant sets occur becuse researchers varied experimental conditions such as host plant, humidity and daylength, or because data were reported for males and females separately. We considered multiple measurements reported in the same publication that were obtained at the same temperature, of the same ontogenetic stage of individuals of the same population (sp and locality) to be redundant, and we obtained the geometric mean of all treatments (host, sex, daylength, humidity,etc) at the same temperature for analyses. 
Consolidated sets kept the number (unique set identifier) of the first set they include.


```{r}
# make a new column that unifies redundant sets
Ana_survival <- Ana_survival %>% 
  mutate(redundancy_id = as.numeric(factor(paste(sp, locality, lifestage, id))))


mean_redundant <- function(dat){
newset <- dat$set [[1]]
dat %>%
  select(temp, sp, survival) %>% 
  group_by(temp) %>% 
  summarise(survival_c = geometric.mean(survival),
            survival2 = mean(survival), 
            newset = newset,
            n_sets = length(survival),.groups = "drop") 
}

# Apply function to Ana_survival

Ana_survival_consolidated <- Ana_survival %>% 
  group_by(redundancy_id) %>% 
  nest() %>% 
  mutate(consolidated = map(data, mean_redundant)) %>% 
  select(redundancy_id, consolidated) %>% 
  unnest(col = consolidated)# Function

# re-name variables. Note that consolidated sets summarize multiple sets and got assigned the number (set id value) of the first set.
names(Ana_survival_consolidated) <- c("redundancy_id", "temp", "survival", "survival2", "set", "n_sets")
```


## 4.Linear interpolation 

Obtain predicted survival in the range from 0 to 40C using linear interpolation. 

```{r}
#create curves: 1- Make interpolation function:
interpolate <- function(dat) {
  out <- seq(from = 0, to = 40, by = 0.1)
  as.data.frame(approx(x = dat$temp, y = dat$survival, xout = out, method = "linear", rule = 1))
  
}

Predicted_Ana_survival <-  Ana_survival_consolidated %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(predicted = map(data,interpolate)) %>% 
  select(set, predicted) %>% 
  unnest(cols = predicted)
names(Predicted_Ana_survival) <- c("set", "temp", "survival")
```



## 5. Calculate Sl, Sh, Sbreadth

```{r}

get_interval <- function(dat){
  dat <- dat[!is.na(dat$survival) & !is.na(dat$temp),] # Remove treatments with no survival data
  Smin <- min(dat$temp)
  Smax <- max(dat$temp)
  y_max <- max(dat$survival) # get maximum value for survival
  x_max <- dat[dat$survival == y_max, "temp"][[1]][[1]]# get temperature that maximizes survival
  cold <- filter(dat, temp < x_max)  
  hot <- filter(dat, temp > x_max)  
  limit <- y_max/2
  if (min(cold$survival) > limit){
    print("cold incomplete")
    y_mincold <- NA
    x_min_cold <- NA
    grade_cold <- "incomplete"
  } else{
    y_mincold <- min(filter(cold, survival >= y_max/2)[["survival"]])
    x_min_cold <- min(filter(cold, survival >= y_max/2)[["temp"]])
    grade_cold <- "complete"
  }
  
  if(min(hot$survival) > limit){
    print("heat incomplete")
    y_minhot <- NA
    x_min_hot <- NA 
    grade_hot <- "incomplete"
  } else {
    y_minhot <- min(filter(hot, survival >= y_max/2)[["survival"]])
    x_min_hot <- max(filter(hot, survival >= y_max/2)[["temp"]])
    grade_hot <- "complete"
  }
  
  
  output <- tibble(Smin = Smin,
                   Smax = Smax, 
                   Sl = x_min_cold, 
                   Sh = x_min_hot, 
                   Sbreadth = Sh - Sl, 
                   Sgradecold = grade_cold, 
                   Sgradehot = grade_hot)
  
  
}


```


```{r}

# apply function to each set

Interval <- Predicted_Ana_survival %>% 
  group_by(set) %>% 
  nest() %>% 
  mutate(interval = map(data,get_interval)) %>% 
  select(set, interval) %>% 
  unnest(cols = interval)

add_info <- Ana_survival %>% 
  select(set, sp, family, lifestage, lat, lon, locality, quality) %>% 
  distinct()
Interval_table1 <- inner_join(Interval, add_info, by = "set")
Interval_table1$lifestage <- factor(Interval_table1$lifestage, levels = c("egg", "larva", "pupa", "eggtoemergence"))
```


```{r}
# write_csv(Interval_table1,"~/Desktop/Projects/ThermalTManuscriptScripts/SurvivalTable.csv")
```


