---
title: "Analyze survival limits"
output:
  html_document:
    df_print: paged
---
## Contents
1. Import data and housekeeping

2. Sl
2.1  Model selection
2.2  Implement phylogenetic correction

3. Sh
3.1  Model selection
3.2  Phylogenetic correction

4. Sbreadth
4.1  Model selection
4.2  Implement phylogenetic correction


# 1. Import data and housekeeping
```{r}
 
#library(INLA) 
# install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
library(tidyverse)
library(readxl)
library(cowplot)
library(psych) # to get geometric mean
library(lme4)
library(car)
library("dotwhisker")  # to make coefficient plots
library(sjPlot) #prediction plots
library(sjmisc)
library(broom)
library(effects)
library(ggeffects)
library(Hmisc)
library(phyr)
library(ape)
library(nlme)
library(stringr)
library(phylotools)
library(phytools)
library(rr2)
# sessionInfo()
```


```{r}
responses <- read_csv("~/Desktop/Projects/ThermalTManuscriptScripts/SurvivalTable.csv")
tree <- read.tree("~/Desktop/Projects/ThermalStressCompilation/GTRG_Unpartitioned_constrain.raxml.bestTree.tre")
unique(responses$lifestage)
```

We discarded "eggtoemergence" category in the subsequent analyses because it is a combination of the other three. We also discarded data from sets that were a combination of localities "combination".

```{r}
# convert character to factor
responses <- responses %>%
  mutate_if(is.character, factor)
# discard non-pertinent data

responses_table <- responses %>% 
  filter(lifestage == "egg"| lifestage == "larva"| lifestage == "pupa", 
         #quality!= "inferred", 
         quality != "combination") %>% 
  mutate(absLat = abs(lat)) 
responses_table$lifestage <- factor(responses_table$lifestage)
responses_table$sp <- factor(responses_table$sp)

# Select variables to include in analyses
S.set <- select(responses_table, set,  Sl, Sh, Sbreadth,lifestage, absLat, sp, family, quality)
S.set$genus <- word(S.set$sp, 1)
S.set$species <- word(S.set$sp, 2)
#filter(responses, set == 1005)
# save a list of sets included in analyses
S.sets <- S.set %>% 
  select(set)

```
There were `length(unique(S.set$sp))`species
There were `length(unique(S.set$set))`sets
There were `nrow(filter(S.set, quality == "inferred")["quality"])/length(unique(S.set$set))`species
41/173
# 2. Sl
Sl ranged from `r min(S.set$Sl, na.rm = T)` to `r max(S.set$Sl, na.rm = T)`, mean =  `r mean(S.set$Sl, na.rm = T)`, SD = `r sd(S.set$Sl, na.rm = T)`
Latitude ranged from `r min(S.set$absLat, na.rm = T)` to `r max(S.set$absLat, na.rm = T)`

## 2.1 Model selection

```{r}
DataLow <- filter(S.set, !is.na(Sl))
Sl_i <- lmer(Sl ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(DataLow)) #
Sl_a <- lmer(Sl ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(DataLow)) #

anova(Sl_i, Sl_a, test.statistic = "Chisq") 
```

```{r}
Sl_i_ML <- lmer(Sl ~ -1 + absLat + lifestage + lifestage:absLat + (1|sp), REML = F, data = as.data.frame(DataLow)) #
summary(Sl_i_ML)

```


```{r}
# Predictions plot
DataLow$group <- factor(DataLow$lifestage)
preSlow <- ggpredict(Sl_i, terms = c("absLat","lifestage"))
# write_csv(preSlow,"~/Desktop/Projects/ThermalTManuscriptScripts/preSlow.csv")

Slow_plot <- ggplot(preSlow, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(S[l]))+
  xlab("Absolute latitude")+
  ylim(0, 40)+
  annotate("text", x = 0, y = 40, label = "G", size = 5)+
  geom_point(data = DataLow, mapping = aes(x = absLat, y = Sl, col = group), alpha = 0.6, shape = 16)+
  #geom_point(data = S.set, mapping = aes(x = absLat, y = Dmin, col = group), alpha = 1, shape = 22)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")

Slow_plot
```
 
# Phylogenetic correction
Change species' names of the following species to reflect substitutions due to lack of data and to correct mispellings

Substitutions:
 In tree                  In dataset    
Ephestia columbiella for     Ephestia calidella #### Not in analyses because source population was "inferred"
Episimus tyrius for          Episimus utilis 
Euzopherodes allocrossa for Euzopherodes vapidella ### No development rate data available
Marmara arbutiella for      Marmara gulosa
Protodeltote albidula for   Naranga aenescens

Mispellings: 
Hyphantria cunea instead of   Hypantria cunea
Ameyelois transitella instead of Amyelois transitella,
```{r}

DataLow$original_species <- DataLow$sp
DataLow$sp <- ifelse(DataLow$original_species == "Episimus utilis", "Episimus tyrius", ifelse(DataLow$original_species == "Marmara gulosa", "Marmara arbutiella", ifelse(DataLow$original_species == "Naranga aenescens", "Protodeltote albidula", ifelse(DataLow$original_species == "Hypantria cunea", "Hyphantria cunea",  ifelse(DataLow$original_species == "Ameyelois transitella", "Amyelois transitella", ifelse(DataLow$original_species == "Utethesia ornatrix", "Utetheisa ornatrix",              ifelse(DataLow$original_species == "Euzopherodes vapidella", "Euzopherodes allocrossa", ifelse(DataLow$original_species == "Ephestia calidella","Ephestia columbiella", as.character(DataLow$original_species)))))))))

Sspecies <- tibble(sp = unique(DataLow$sp))
tree_species <-  tibble(sp = as.character(paste(word(tree$tip.label, sep = "_", 4), word(tree$tip.label, sep = "_", 5))))
V_species <- unique(DataLow$sp)
# setdiff(Sspecies, tree_species)
# all species in DataLow are in the tree. 

# setdiff(tree_species, Sspecies) # the tree has 64 species more than S.set

```
Trim the tree so it only includes the species in the P.set. First the tips have to be relabeled so they include only the species names
```{r}
rename <- as.data.frame(tibble(oldname = tree$tip.label, 
                 newname = tree_species$sp))
tree_relabel <- sub.taxa.label(tree, rename)
Sl.tree <- keep.tip(tree_relabel, V_species)
plotTree(Sl.tree)
```


```{r}
Slow_phylo <- pglmm(Sl ~ -1 + lifestage + absLat + (1|sp__), tree = tree, bayes= F, data = DataLow, cov_ranef = list(sp = Sl.tree), REML = F)

Slow_nophylo <- pglmm(Sl ~ -1 + lifestage + absLat + (1|sp), tree = tree, bayes= F, data = DataLow, cov_ranef = list(sp = Sl.tree), REML = F)

```


```{r}
summary(Slow_phylo)
```

```{r}
summary(Slow_phylo)
```


Asses the effect of phylogeny on Slow
```{r}

#R2(Slow_phylo, Slow_nophylo) # if this does not work unload and reload rr2
R2.lik(Slow_phylo, Slow_nophylo) # if this does not work unload and reload rr2


```

#3. Shigh 
Shigh ranged from `r min(S.set$Sh, na.rm = T)` to `r max(S.set$Sh, na.rm = T)`, mean =  `r mean(S.set$Sh, na.rm = T)`, SD = `r sd(S.set$Sh, na.rm = T)`
Latitude ranged from `r min(S.set$absLat, na.rm = T)` to `r max(S.set$absLat, na.rm = T)`

```{r}

DataHigh <- filter(S.set, !is.na(Sh))

```


## 3.1 Model selection

```{r}
Sh_i <- lmer(Sh ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(DataHigh)) #
Sh_a <- lmer(Sh ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(DataHigh)) #

anova(Sh_i, Sh_a, test.statistic = "Chisq")
```

```{r}
Sh_a_ML <- lmer(Sh ~ -1 + absLat + lifestage + (1|sp), REML = F, data = as.data.frame(DataHigh)) #
summary(Sh_a_ML)
```

```{r}
# Predictions plot

DataHigh$group <- factor(DataHigh$lifestage)
preSu <- ggpredict(Sh_i, terms = c("absLat","lifestage"))

#write_csv(preSu,"~/Desktop/Projects/ThermalTManuscriptScripts/preSu.csv")
Supper_plot <- ggplot(preSu, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  ylab(expression(S[h]))+
  xlab("Absolute latitude")+
  ylim(10, 50)+
  #annotate("text", x = 0, y = 40, label = "X", size = 5)+
  geom_point(data = DataHigh, mapping = aes(x = absLat, y = Sh, col = group), alpha = 0.6, shape = 1)+
  #geom_point(data = S.set, mapping = aes(x = absLat, y = Dmax, col = group), alpha = 0.6, shape = 22)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")

Supper_plot
```

## 3.2 Phylogenetic correction

### 3.2.1 Adjust data table 
Change species' names of the following species to reflect substitutions due to lack of data and to correct misspellings
Substitutions:
 In tree                  In dataset    
Ephestia columbiella for     Ephestia calidella #### Not in analyses because source population was "inferred"
Episimus tyrius for          Episimus utilis 
Euzopherodes allocrossa for Euzopherodes vapidella ### No development rate data available
Marmara arbutiella for      Marmara gulosa
Protodeltote albidula for   Naranga aenescens

Mispellings: 
Hyphantria cunea instead of   Hypantria cunea
Ameyelois transitella instead of Amyelois transitella,



```{r}
DataHigh$original_species <- DataHigh$sp
DataHigh$sp <- ifelse(DataHigh$original_species == "Episimus utilis", "Episimus tyrius", ifelse(DataHigh$original_species == "Marmara gulosa", "Marmara arbutiella", ifelse(DataHigh$original_species == "Naranga aenescens", "Protodeltote albidula", ifelse(DataHigh$original_species == "Hypantria cunea", "Hyphantria cunea",  ifelse(DataHigh$original_species == "Ameyelois transitella", "Amyelois transitella", ifelse(DataHigh$original_species == "Utethesia ornatrix", "Utetheisa ornatrix", ifelse(DataHigh$original_species == "Ephestia calidella","Ephestia columbiella", as.character(DataHigh$original_species))))))))

Dspecies <- tibble(sp = unique(DataHigh$sp))
V_species <- unique(DataHigh$sp)
tree_species <-  tibble(sp = as.character(paste(word(tree$tip.label, sep = "_", 4), word(tree$tip.label, sep = "_", 5))))

#setdiff(Dspecies, tree_species)
# all species in DataHigh are in the tree. 

#setdiff(tree_species, Dspecies) # the tree has 45 species more than DataHigh
```


Trim the tree so it only includes the species in DataHigh. First the tips have to be relabeled so they include only the species names
```{r}
rename <- as.data.frame(tibble(oldname = tree$tip.label, 
                 newname = tree_species$sp))
tree_relabel <- sub.taxa.label(tree, rename)
Sh.tree <- keep.tip(tree_relabel, V_species)
plotTree(Sh.tree)
```





```{r}
Shigh_phylo <- pglmm(Sh ~ -1 + lifestage + absLat + (1|sp__), tree = tree, bayes= F, data = DataHigh, cov_ranef = list(sp = Sh.tree), REML = F)

Shigh_nophylo <- pglmm(Sh ~ -1 + lifestage + absLat + (1|sp), tree = tree, bayes= F, data = DataHigh, cov_ranef = list(sp = Sh.tree), REML = F)

```


```{r}
Shigh_phylo
```
```{r}
Shigh_phylo
```



Asses the effect of phylogeny
```{r}
R2.lik(Shigh_phylo, Shigh_nophylo) # if this does not work unload and reload rr2
```




# 4. Sbreadth
Sbreadth ranged from `r min(S.set$Sbreadth, na.rm = T)` to `r max(S.set$Sbreadth, na.rm = T)`, mean =  `r mean(S.set$Sbreadth, na.rm = T)`, SD = `r sd(S.set$Sbreadth, na.rm = T)`
Latitude ranged from `r min(S.set$absLat, na.rm = T)` to `r max(S.set$absLat, na.rm = T)`

## 3.1 Model selection

```{r}
DataB <- filter(S.set, !is.na(Sbreadth))
Sbreadth_i <- lmer(Sbreadth ~ -1 + lifestage + absLat + lifestage:absLat + (1|sp), data = as.data.frame(DataB)) #
Sbreadth_a <- lmer(Sbreadth ~ -1 + absLat + lifestage + (1|sp), data = as.data.frame(DataB)) #

anova(Sbreadth_i, Sbreadth_a, test.statistic = "Chisq") 
```

```{r}
Sbreadth_a_ML <- lmer(Sbreadth ~ -1 + absLat + lifestage + (1|sp), REML = F, data = as.data.frame(DataB)) #
summary(Sbreadth_a_ML)
```


```{r}
# Predictions plot
DataB$group <- factor(DataB$lifestage)
presw <- ggpredict(Sbreadth_i, terms = c("absLat","lifestage"))
#write_csv(presw,"~/Desktop/Projects/ThermalTManuscriptScripts/presw.csv")

Sbreadth_plot <- ggplot(presw, aes(x = x, y = predicted, colour = group, fill = group)) +
  geom_line()+
   geom_ribbon(aes(ymin = conf.low, ymax = conf.high, col = group), alpha = 0.1)+
  scale_fill_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_color_viridis_d(option = "plasma", begin = 0, end = 0.8)+
  scale_shape_manual(values = c(1,16,8))+
  ylab(expression(S[breadth]))+
  xlab("Absolute latitude")+
  ylim(0, 40)+
  annotate("text", x = 0, y = 40, label = "X", size = 5)+
  geom_point(data = DataB, mapping = aes(x = absLat, y = Sbreadth, col = group), alpha = 0.6, shape = 0)+
  theme_cowplot()+
  theme(legend.title = element_blank(), legend.position = "none")

Sbreadth_plot
```

## 4.2 Phylogenetic correction


Change species' names of the following species to reflect substitutions due to lack of data and to correct misspellings
Substitutions:
 In tree                  In dataset    
Ephestia columbiella for     Ephestia calidella #### Not in analyses because source population was "inferred"
Episimus tyrius for          Episimus utilis 
Euzopherodes allocrossa for Euzopherodes vapidella ### No development rate data available
Marmara arbutiella for      Marmara gulosa
Protodeltote albidula for   Naranga aenescens

Mispellings: 
Hyphantria cunea instead of   Hypantria cunea
Ameyelois transitella instead of Amyelois transitella,



```{r}
DataB$original_species <- DataB$sp
DataB$sp <- ifelse(DataB$original_species == "Episimus utilis", "Episimus tyrius", ifelse(DataB$original_species == "Marmara gulosa", "Marmara arbutiella", ifelse(DataB$original_species == "Naranga aenescens", "Protodeltote albidula", ifelse(DataB$original_species == "Hypantria cunea", "Hyphantria cunea",  ifelse(DataB$original_species == "Ameyelois transitella", "Amyelois transitella", ifelse(DataB$original_species == "Utethesia ornatrix", "Utetheisa ornatrix", ifelse(DataB$original_species == "Ephestia calidella","Ephestia columbiella", as.character(DataB$original_species))))))))

Dspecies <- tibble(sp = unique(DataB$sp))
V_species <- unique(DataB$sp)
tree_species <-  tibble(sp = as.character(paste(word(tree$tip.label, sep = "_", 4), word(tree$tip.label, sep = "_", 5))))

#setdiff(Dspecies, tree_species)
# all species in DataB are in the tree. 

#setdiff(tree_species, Dspecies) # the tree has 45 species more than DataB
```


Trim the tree so it only includes the species in DataHigh. First the tips have to be relabeled so they include only the species names
```{r}
rename <- as.data.frame(tibble(oldname = tree$tip.label, 
                 newname = tree_species$sp))
tree_relabel <- sub.taxa.label(tree, rename)
Sb.tree <- keep.tip(tree_relabel, V_species)
plotTree(Sb.tree)
```






Fit model
```{r}

Sbreadth_phylo <- pglmm(Sbreadth ~ -1 + lifestage + absLat + (1|sp__), tree = tree, bayes= F, data = DataB, cov_ranef = list(sp = Sb.tree), REML = F)

Sbreadth_nophylo <- pglmm(Sbreadth ~ -1 + lifestage + absLat + (1|sp), tree = tree, bayes= F, data = DataB, cov_ranef = list(sp = Sb.tree), REML = F)

```


```{r}
Sbreadth_phylo
```




```{r}
Sbreadth_phylo
```


Asses the effect of phylogeny
```{r}
 R2.lik(Sbreadth_phylo,Sbreadth_nophylo) # if this does not work unload and reload rr2

```





